---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.  -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# india.air

<!-- badges: start -->
<!-- badges: end -->

The goal of `india.air` is to make Indian air pollution data easily available for analysis and visualization in R.

```{r, message=FALSE, echo=FALSE}
library(india.air)
library(tidyverse)
library(lubridate)
library(RColorBrewer)
library(viridis)
```


```{r, echo=FALSE}
# wranngling
three_cities <- india_air %>%
  filter(city %in% c("Mumbai",
                     "Delhi",
                     "Chennai"))
```


```{r, message=FALSE, echo=FALSE, warning=FALSE, fig.retina = 3}
ggplot(data = three_cities, mapping = aes(x = date,
                                          y = PM2.5,
                                          color = city)) +
  geom_line(size = 0.7,
            alpha = 0.7) +
  xlim(mdy("1/1/2017"), mdy("12/31/2019")) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.title.x = element_blank(),
        legend.box.margin = margin(0, 0, 0, 0, "mm")) +
  scale_y_continuous(limits = c(0, 610),
                     expand = c(0, 0),
                     name = "PM2.5 (\u03BCg/m\u00b3)") +
  scale_color_brewer(palette = "Accent") +
  labs(color = "City",
       title = "PM2.5 levels in three Indian cities, 2017-2019")
```


## Installation

The development version of `india.air` is available from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("Reed-Math241/pkgGrpr")
```

## About the data

This data was made available by the [Central Pollution Control Board of India](https://cpcb.nic.in/) and compiled by [Vopani on kaggle](https://www.kaggle.com/rohanrao/air-quality-data-in-india).

The data is within the public domain under the CC-0 license.

The package `india.air` contains one dataset, `india_air`. See `?india_air` for a description of each of the variables.

```{r}
head(india_air)
```

The dataset contains air pollution data for six Indian cities between the years of 2015 and 2020 at the day level. Not all pollution measurements were collected in each city on each date, so there is some intrinsic missingness within the dataset.


## Examples

`india.air` is great for visualizing patterns in air pollution over time. For example,

```{r, message=FALSE, echo=FALSE, warning=FALSE, fig.retina = 3}
india_air %>%
  filter(city == "Delhi") %>%
ggplot(mapping = aes(x = date,
                     y = NOx)) +
  geom_line(size = 0.7,
            alpha = 0.9,
            color = "turquoise") +
  theme_minimal() +
  theme(legend.position = "top",
        axis.title.x = element_blank()) +
  scale_y_continuous(limits = c(0, 210),
                     expand = c(0, 0),
                     name = "[NOx] (\u03BCg/m\u00b3)") +
  scale_x_date(limits = c(mdy("1/1/2017"), mdy("12/31/2019"))) +
  labs(title = "NOx concentrations in Delhi, 2017-2019")
```

```{r, message=FALSE, echo=FALSE, warning=FALSE, fig.retina = 3}
india_air %>%
  filter(city == "Ahmedabad",
         date >= mdy("1/1/2018") & date <= mdy("12/30/2018")) %>%
ggplot(mapping = aes(x = date,
                     y = AQI)) +
  annotate("rect", xmin = mdy("12/31/2017"),      #AQI bucket "good"
           xmax = mdy("12/31/2018"),
           ymin = 0, ymax = 50,
           alpha = 0.5, fill = "#009900") +
  annotate("rect", xmin = mdy("12/31/2017"),      #AQI bucket "satisfactory"
           xmax = mdy("12/31/2018"),
           ymin = 50.1, ymax = 100,
           alpha = 0.5, fill = "#669900") +
  annotate("rect", xmin = mdy("12/31/2017"),      #AQI bucket "moderately polluted"
           xmax = mdy("12/31/2018"),
           ymin = 100.1, ymax = 200,
           alpha = 0.5, fill = "#CC9900") +
  annotate("rect", xmin = mdy("12/31/2017"),      #AQI bucket "poor"
           xmax = mdy("12/31/2018"),
           ymin = 200.1, ymax = 300,
           alpha = 0.5, fill = "#996600") +
  annotate("rect", xmin = mdy("12/31/2017"),      #AQI bucket "very poor"
           xmax = mdy("12/31/2018"),
           ymin = 300.1, ymax = 400,
           alpha = 0.5, fill = "#663300") +
  annotate("rect", xmin = mdy("12/31/2017"),      #AQI bucket "severe"
           xmax = mdy("12/31/2018"),
           ymin = 400.1, ymax = 2070,
           alpha = 0.5, fill = "#330000") +
  geom_line(size = 0.7,
            alpha = 0.9,
            color = "black") +
  theme_minimal() +
  theme(legend.position = "top",
        axis.title.x = element_blank()) +
  scale_y_continuous(expand = c(0, 0),
                     name = "AQI") +
  scale_x_date(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  labs(title = "AQI in Ahmedabad, 2018") +
  annotate("rect", xmin = mdy("7/9/2018"),                #background of key
           xmax = mdy("10/13/2018"),
           ymin = 1080, ymax = 1800,
           fill = "white") +
  annotate("text", x = mdy("8/27/2018"),
           y = 1750, label = "AQI Bucket",
           color = "grey20") +
  annotate("rect", xmin = mdy("7/13/2018"),               #good
           xmax = mdy("7/22/2018"),
           ymin = 1610, ymax = 1680,
           fill = "#009900", alpha = 0.5) +
  annotate("text", x = mdy("8/7/2018"),
           y = 1650, label = "Good", size = 3.5,
           color = "grey20", family = "sans") +
  annotate("rect", xmin = mdy("7/13/2018"),               #satisfactory
           xmax = mdy("7/22/2018"),
           ymin = 1510, ymax = 1580,
           fill = "#669900", alpha = 0.5) +
  annotate("text", x = mdy("8/20/2018"),
           y = 1550, label = "Satisfactory", size = 3.5,
           color = "grey20", family = "sans") +
  annotate("rect", xmin = mdy("7/13/2018"),               #moderately polluted
           xmax = mdy("7/22/2018"),
           ymin = 1410, ymax = 1480,
           fill = "#CC9900", alpha = 0.5) +
  annotate("text", x = mdy("9/3/2018"),
           y = 1450, label = "Moderately polluted", size = 3.5,
           color = "grey20", family = "sans") +
  annotate("rect", xmin = mdy("7/13/2018"),               #poor
           xmax = mdy("7/22/2018"),
           ymin = 1310, ymax = 1380,
           fill = "#996600", alpha = 0.5) +
  annotate("text", x = mdy("8/7/2018"),
           y = 1350, label = "Poor", size = 3.5,
           color = "grey20", family = "sans") +
  annotate("rect", xmin = mdy("7/13/2018"),               #very poor
           xmax = mdy("7/22/2018"),
           ymin = 1210, ymax = 1280,
           fill = "#663300", alpha = 0.5) +
  annotate("text", x = mdy("8/15/2018"),
           y = 1250, label = "Very poor", size = 3.5,
           color = "grey20", family = "sans") +
  annotate("rect", xmin = mdy("7/13/2018"),               #severe
           xmax = mdy("7/22/2018"),
           ymin = 1110, ymax = 1180,
           fill = "#330000", alpha = 0.5) +
  annotate("text", x = mdy("8/11/2018"),
           y = 1150, label = "Severe", size = 3.5,
           color = "grey20", family = "sans")
```


`india.air` is also great for comparing pollution among cities:

```{r, message=FALSE, echo=FALSE, warning=FALSE}
ggplot(data = india_air, mapping = aes(x = date,
                                       y = PM2.5,
                                       color = city)) +
  geom_line(size = 0.7,
            alpha = 0.7) +
  xlim(mdy("12/1/2019"), mdy("12/31/2019")) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.title.x = element_blank()) +
  scale_y_continuous(limits = c(0, 350),
                     expand = c(0, 0)) +
  scale_color_brewer(palette = "Dark2") +
  labs(color = "City",
       title = "PM2.5 levels in six Indian cities, December 2019",
       y = "PM2.5 (\u03BCg/m\u00b3)")

#micrograms per meter cubed: \u03BCg/m\u00b3
#subscript number for chemical formulas : \u208_ , replace underscore with desired number
```






